---
title: "Cleaning NTIS Data: Before and After ntistools"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Cleaning NTIS Data: Before and After ntistools}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Why ntistools?

NTIS data-cleaning scripts tend to repeat the same `case_when()` patterns over
and over: combining binary indicators, removing sentinels, collapsing Likert
scales, labeling columns, and so on. A typical script might contain 20+ blocks
of nearly identical code like this:

```r
# This same pattern, repeated for every binary variable...
ProgDem_Veterans = case_when(
  ProgDem_Veterans == 1 ~ "Serving veterans",
  ProgDem_Veterans == 0 ~ "Not serving veterans",
  .default = NA_character_
)
```

**ntistools** replaces each pattern with a single, purpose-built function so
your scripts are shorter, easier to read, and less error-prone. This vignette
walks through each function, showing the verbose "before" code alongside the
concise "after" equivalent, all using a small synthetic dataset.

## Setup

We start by creating a synthetic dataset that mimics the structure of real NTIS
survey data. Each column represents a common pattern you'll encounter:

```{r setup, message = FALSE}
library(ntistools)
library(dplyr)

survey <- data.frame(
  # Binary geographic indicators (0 = no, 1 = yes)
  GeoAreas_Local          = c(1, 0, 0, NA, 0, 1),
  GeoAreas_MultipleLocal  = c(0, 0, 1, NA, 0, 0),
  GeoAreas_RegionalWithin = c(0, 0, 0, NA, 0, 0),
  GeoAreas_MultipleState  = c(0, NA, 1, NA, 0, 0),
  GeoAreas_RegionalAcross = c(0, NA, 0, NA, 0, 1),

  # Disruption indicators (0 = no, 1 = yes)
  LLLost  = c(1, 0, 1, NA, 0, 1),
  LLDelay = c(1, 1, 0, NA, 0, 1),
  LLStop  = c(0, 0, 0, NA, 0, 1),

  # ProgDem-style columns (0 = no, 1 = yes primary, 2 = yes secondary, 98 = NA)
  ProgDem_Children = c(0, 1, 2, 98, 0, 1),
  ProgDem_Elders   = c(1, 0, 98, 2, 0, 1),

  # Sentinel values (98 = "don't know")
  DonImportance = c(3, 1, 98, 2, 98, 4),

  # Flag-based imputation columns
  # PplSrv_NumWait_NA_X = 1 means "respondent confirmed they had zero people
  # waiting", so NA should be replaced with 0
  PplSrv_NumWait      = c(50, NA, 10, NA, 30, NA),
  PplSrv_NumWait_NA_X = c(0,   1,  0,  0,  0,  1),

  # Imputation with year-suffixed variable names
  # The variable has a year suffix but the flag column does not
  Staff_RegVlntr_2023 = c(NA, 5,  NA, 10, NA, 3),
  Staff_RegVlntr_2024 = c(2,  NA, NA, 7,  NA, 4),
  Staff_RegVlntr_NA   = c(1,  0,  0,  0,  1,  0),

  # Likert scales (1-5, with 97 = "not applicable")
  FinanceChng_Benefits  = c(1, 2, 3, 4, 5, 97),
  FinanceChng_Salaries  = c(5, 4, 3, 2, 1, 97),
  FinanceChng_TotExp    = c(2, 3, 97, 5, 1, 4),

  # Finance change indicators (1 = yes, 0 = no, 97 = not applicable)
  FinanceChng_Reserves = c(1, 0, 97, 1, 0, 1),
  PrgSrvc_Suspend      = c(1, 0, 97, 1, 0, 0),

  # Parent/child regulation pattern
  Regulations         = c(1, 0, 97, 1, 0, 1),
  Regulations_Federal = c(1, NA, NA, 0, NA, 1),
  Regulations_State   = c(0, NA, NA, 1, NA, 1),

  # Staff filter columns
  HaveStaff      = c(1, 1, 0, 1, 0, 1),
  StaffVacancies = c(3, 1, 2, 0, 4, 1),
  BenefitsImpact = c(2, 3, 1, 4, 2, 3)
)
```

---

## 1. `combine_binary()` --- Grouping binary indicators

**The problem:** You have several related 0/1 columns and need a single column
that is 1 if **any** of them are 1. For example, in the NTIS extraction script,
`GeoAreas_Local`, `GeoAreas_MultipleLocal`, and `GeoAreas_RegionalWithin` are
combined into `GeoAreas_ServedLocal`.

### Before

```{r}
before <- survey %>%
  mutate(
    GeoAreas_Locally = case_when(
      GeoAreas_Local == 1 | GeoAreas_MultipleLocal == 1 |
        GeoAreas_RegionalWithin == 1 ~ 1,
      is.na(GeoAreas_Local) & is.na(GeoAreas_MultipleLocal) &
        is.na(GeoAreas_RegionalWithin) ~ NA,
      .default = 0
    )
  )

before %>% select(starts_with("GeoAreas_L"), GeoAreas_MultipleLocal,
                  GeoAreas_RegionalWithin)
```

### After

```{r}
after <- survey %>%
  combine_binary(GeoAreas_Locally,
                 GeoAreas_Local, GeoAreas_MultipleLocal,
                 GeoAreas_RegionalWithin)

after %>% select(starts_with("GeoAreas_L"), GeoAreas_MultipleLocal,
                 GeoAreas_RegionalWithin)
```

### Option: `strict_na` for stricter NA handling

By default, `combine_binary()` only returns NA when **all** source columns are
NA. If a row has `c(0, NA)`, the result is 0.

Sometimes you need stricter behavior: if **any** source is NA (and none are 1),
the result should be NA. This matches the `binary_flag()` function used in older
NTIS scripts. Set `strict_na = TRUE` to get this behavior.

```{r}
# Data where one source is 0 and the other is NA
strict_example <- data.frame(
  a = c(0,  1,  NA, 0),
  b = c(NA, NA, NA, 0)
)

# Default: lenient (0 + NA = 0)
combine_binary(strict_example, result_lenient, a, b) %>%
  select(a, b, result_lenient)

# Strict: any NA poisons the result (0 + NA = NA), unless a 1 is present
combine_binary(strict_example, result_strict, a, b, strict_na = TRUE) %>%
  select(a, b, result_strict)
```

**When to use `strict_na = TRUE`:** Use it when a missing value means you
genuinely don't know if the respondent would have answered "yes." In the NTIS
extraction script, the `GeoAreas_ServedLocal` and `GeoAreas_Servedmultistate`
variables use strict mode.

---

## 2. `count_binary()` --- Any / count / all from binary indicators

Creates three summary columns at once from a set of binary columns:

- `{prefix}_any`: 1 if **any** source is 1
- `{prefix}_count`: how many sources are 1
- `{prefix}_all`: 1 if **every** source is 1

All three are NA when every source column is NA.

### Before

```{r}
before <- survey %>%
  mutate(
    LLDisruption_any = case_when(
      LLLost == 1 | LLDelay == 1 | LLStop == 1 ~ 1,
      is.na(LLLost) & is.na(LLDelay) & is.na(LLStop) ~ NA,
      .default = 0
    ),
    LLDisruption_count = case_when(
      is.na(LLLost) & is.na(LLDelay) & is.na(LLStop) ~ NA_real_,
      .default = rowSums(across(c(LLLost, LLDelay, LLStop),
                                ~ . == 1), na.rm = TRUE)
    ),
    LLDisruption_all = case_when(
      LLLost == 1 & LLDelay == 1 & LLStop == 1 ~ 1,
      is.na(LLLost) & is.na(LLDelay) & is.na(LLStop) ~ NA,
      .default = 0
    )
  )

before %>% select(starts_with("LL"))
```

### After

```{r}
after <- survey %>%
  count_binary("LLDisruption", LLLost, LLDelay, LLStop)

after %>% select(starts_with("LL"))
```

**Tip:** If you only need the `_any` column (not `_count` or `_all`), use
`combine_binary()` instead --- it creates a single column and lets you name it
whatever you want.

---

## 3. `recode_binary()` --- Collapse multi-level to 0/1

**The problem:** NTIS demographic columns use 0 = "no", 1 = "yes (primary)",
2 = "yes (secondary)", and 98 = "don't know." You need to collapse these to
simple 0/1. In the extraction script, all `ProgDem_` columns go through this
step.

### Before

```{r}
recode_progdem <- function(x) {
  case_when(
    x == 0 ~ 0, x == 1 ~ 1, x == 2 ~ 1,
    x == 98 ~ NA_real_, TRUE ~ NA_real_
  )
}

before <- survey %>%
  mutate(across(c(ProgDem_Children, ProgDem_Elders),
                recode_progdem))

before %>% select(starts_with("ProgDem"))
```

### After

```{r}
after <- survey %>%
  recode_binary(c("ProgDem_Children", "ProgDem_Elders"))

after %>% select(starts_with("ProgDem"))
```

### Customizing value mappings

The defaults (`ones = c(1, 2)`, `zeros = 0`, `na_values = 98`) work for
`ProgDem_` columns, but you can customize them for other patterns:

```{r}
# Example: a column where 1-3 mean "yes" and 4-5 mean "no"
custom_example <- data.frame(rating = c(1, 2, 3, 4, 5, 98))

recode_binary(custom_example, "rating",
              ones = c(1, 2, 3),
              zeros = c(4, 5),
              na_values = 98) %>%
  pull(rating)
```

---

## 4. `recode_sentinel()` --- Remove sentinel values

Replace survey sentinel values (e.g., 98 = "don't know") with `NA`. The
extraction script removes 98s from dozens of columns before analysis.

### Before

```{r}
before <- survey %>%
  mutate(across(
    all_of("DonImportance"),
    ~ case_when(. == 98 ~ NA, TRUE ~ .)
  ))

before %>% select(DonImportance)
```

### After

```{r}
after <- survey %>%
  recode_sentinel("DonImportance")

after %>% select(DonImportance)
```

### Removing different sentinel values

By default, `recode_sentinel()` removes 98. You can change which values to
remove:

```{r}
# Remove both 97 and 98
survey %>%
  recode_sentinel("DonImportance", values = c(97, 98)) %>%
  select(DonImportance)
```

---

## 5. `impute_from_flag()` --- Fill NA when a flag says "skip was valid"

**The problem:** When a respondent legitimately skipped a numeric question (e.g.,
"How many people are on your waitlist?" when they have no waitlist), the value is
`NA` but a companion flag column (e.g., `PplSrv_NumWait_NA_X`) is set to 1.
These NAs should be replaced with 0 (not left as missing data).

### Before

```{r}
before <- survey %>%
  mutate(
    PplSrv_NumWait = case_when(
      PplSrv_NumWait_NA_X == 1 & is.na(PplSrv_NumWait) ~ 0,
      .default = PplSrv_NumWait
    )
  )

before %>% select(starts_with("PplSrv"))
```

### After

```{r}
after <- survey %>%
  impute_from_flag("PplSrv_NumWait")

after %>% select(starts_with("PplSrv"))
```

### Option: `flag_map` for non-standard flag column names

Sometimes the variable name includes a year suffix (e.g.,
`Staff_RegVlntr_2023`) but the flag column does not (e.g., `Staff_RegVlntr_NA`).
The default `paste0(var, flag_suffix)` logic would look for
`Staff_RegVlntr_2023_NA_X`, which doesn't exist.

Use `flag_map` to tell the function exactly which flag column to use for each
variable:

```{r}
# Before: six separate case_when blocks in the extraction script
before <- survey %>%
  mutate(
    Staff_RegVlntr_2023 = case_when(
      is.na(Staff_RegVlntr_2023) & Staff_RegVlntr_NA == 1 ~ 0,
      .default = Staff_RegVlntr_2023
    ),
    Staff_RegVlntr_2024 = case_when(
      is.na(Staff_RegVlntr_2024) & Staff_RegVlntr_NA == 1 ~ 0,
      .default = Staff_RegVlntr_2024
    )
  )

before %>% select(starts_with("Staff_RegVlntr"))
```

```{r}
# After: one call handles both year columns
after <- survey %>%
  impute_from_flag(
    vars = c("Staff_RegVlntr_2023", "Staff_RegVlntr_2024"),
    flag_map = c(Staff_RegVlntr_2023 = "Staff_RegVlntr_NA",
                 Staff_RegVlntr_2024 = "Staff_RegVlntr_NA")
  )

after %>% select(starts_with("Staff_RegVlntr"))
```

**Tip:** You can mix `flag_map` and the default suffix in the same call.
Variables listed in `flag_map` use their mapped flag; all others fall back to
`paste0(var, flag_suffix)`.

---

## 6. `collapse_likert()` --- Likert 5-point to 3-point

Collapse a 5-point Likert scale (1--5) into 3 categories (1 = low, 2 = mid,
3 = high), with 97 set to NA. The default `mapping = c(1, 1, 2, 3, 3)` means:

| Original | Maps to | Meaning |
|:--------:|:-------:|---------|
| 1        | 1       | Low (e.g., "Significant decrease") |
| 2        | 1       | Low (e.g., "Some decrease") |
| 3        | 2       | Mid (e.g., "No change") |
| 4        | 3       | High (e.g., "Some increase") |
| 5        | 3       | High (e.g., "Significant increase") |
| 97       | NA      | Not applicable |

### Before

```{r}
before <- survey %>%
  mutate(
    FinanceChng_Benefits = case_match(
      FinanceChng_Benefits,
      1 ~ 1, 2 ~ 1, 3 ~ 2, 4 ~ 3, 5 ~ 3, 97 ~ NA
    )
  )

before %>% select(FinanceChng_Benefits)
```

### After

```{r}
after <- survey %>%
  collapse_likert("FinanceChng_Benefits")

after %>% select(FinanceChng_Benefits)
```

### Applying to multiple columns at once

In the extraction script, many Likert columns are collapsed using the same
mapping. Pass a character vector of column names:

```{r}
likert_cols <- c("FinanceChng_Benefits", "FinanceChng_Salaries",
                 "FinanceChng_TotExp")

after <- survey %>%
  collapse_likert(likert_cols)

after %>% select(all_of(likert_cols))
```

---

## 7. `propagate_parent()` --- Push parent value to children

When a parent question (e.g., "Does your organization face government
regulations?") is 0 ("no") or 97 ("N/A"), the follow-up questions should inherit
that value instead of being left as NA.

### Before

```{r}
before <- survey %>%
  mutate(across(
    c(Regulations_Federal, Regulations_State),
    ~ case_when(
      Regulations == 0 ~ 0,
      Regulations == 97 ~ 97,
      TRUE ~ .
    )
  ))

before %>% select(starts_with("Regulations"))
```

### After

```{r}
after <- survey %>%
  propagate_parent(c("Regulations_Federal", "Regulations_State"),
                   parent_var = "Regulations")

after %>% select(starts_with("Regulations"))
```

### Customizing which parent values propagate

By default, `values = c(0, 97)` (both "no" and "not applicable" propagate). You
can change this:

```{r}
# Only propagate 0 (not 97)
survey %>%
  propagate_parent(c("Regulations_Federal", "Regulations_State"),
                   parent_var = "Regulations",
                   values = 0) %>%
  select(starts_with("Regulations"))
```

---

## 8. `apply_filter()` --- Set variables to NA for ineligible respondents

**The problem:** Some survey questions only apply to certain organizations. For
example, questions about staff vacancies and benefits only make sense for
organizations that have paid staff. For orgs without staff, these values should
be set to NA so they don't affect analysis.

### Before

```{r}
before <- survey %>%
  mutate(across(
    c(StaffVacancies, BenefitsImpact),
    ~ case_when(HaveStaff == 1 ~ ., .default = NA)
  ))

before %>% select(HaveStaff, StaffVacancies, BenefitsImpact)
```

### After

```{r}
after <- survey %>%
  apply_filter(c("StaffVacancies", "BenefitsImpact"),
               condition = HaveStaff == 1)

after %>% select(HaveStaff, StaffVacancies, BenefitsImpact)
```

**Tip:** The `condition` argument accepts any expression you could use in
`dplyr::filter()`. For example, `condition = HaveStaff == 1 & year == "2024"`
would only keep values for orgs with staff in 2024.

---

## 9. `label_binary()` --- Convert 0/1 to descriptive labels

**The problem:** After recoding to 0/1, you often need to convert columns to
descriptive strings for reporting. The NTIS extraction script has 25+ blocks of
identical `case_when()` code like this, one per variable. `label_binary()`
replaces all of them in a single call.

### Before

```{r}
before <- survey %>%
  mutate(
    FinanceChng_Reserves = case_when(
      FinanceChng_Reserves == 1 ~ "Drew on cash reserves",
      FinanceChng_Reserves %in% c(0, 97) ~ "Did not draw on cash reserves",
      .default = NA_character_
    ),
    PrgSrvc_Suspend = case_when(
      PrgSrvc_Suspend == 1 ~ "Paused or suspended services",
      PrgSrvc_Suspend %in% c(0, 97) ~ "Did not pause or suspend services",
      .default = NA_character_
    )
  )

before %>% select(FinanceChng_Reserves, PrgSrvc_Suspend)
```

### After

```{r}
after <- survey %>%
  label_binary(
    labels = list(
      FinanceChng_Reserves = c("Drew on cash reserves",
                               "Did not draw on cash reserves"),
      PrgSrvc_Suspend = c("Paused or suspended services",
                          "Did not pause or suspend services")
    ),
    na_values = 97
  )

after %>% select(FinanceChng_Reserves, PrgSrvc_Suspend)
```

### How `labels` works

The `labels` argument is a named list. Each name is a column in your data, and
each value is a length-2 character vector: `c("label when 1", "label when 0")`.

### Option: `na_values` for sentinel-as-false

Some NTIS variables treat 97 ("not applicable") the same as 0 ("no"). For
example, if an organization answered "not applicable" to whether they drew on
cash reserves, that effectively means they did not. Set `na_values = 97` to map
97 to the false (second) label instead of to `NA`.

```{r}
# Without na_values: 97 becomes NA
survey %>%
  label_binary(
    labels = list(FinanceChng_Reserves = c("Drew", "Did not draw"))
  ) %>%
  select(FinanceChng_Reserves)

# With na_values = 97: 97 maps to the false label
survey %>%
  label_binary(
    labels = list(FinanceChng_Reserves = c("Drew", "Did not draw")),
    na_values = 97
  ) %>%
  select(FinanceChng_Reserves)
```

### Labeling many columns at once

In the extraction script, 16 `ProgDem_` columns are each labeled with a
"Serving..." / "Not serving..." pair. You can handle them all in one call:

```{r}
# First recode to 0/1 (2 = secondary yes), then label
after <- survey %>%
  recode_binary(c("ProgDem_Children", "ProgDem_Elders")) %>%
  label_binary(labels = list(
    ProgDem_Children = c("Serving children and youth",
                         "Not serving children and youth"),
    ProgDem_Elders   = c("Serving seniors",
                         "Not serving seniors")
  ))

after %>% select(starts_with("ProgDem"))
```

### Option: custom `true_values` and `false_values`

For variables that don't use the standard 0/1 coding:

```{r}
# A column where 1,2 = "yes" and 3,4,5 = "no"
age_example <- data.frame(Dem_BChair_Age = c(1, 2, 3, 5, 97, NA))

label_binary(age_example,
  labels = list(Dem_BChair_Age = c("Under 35", "35 or older")),
  true_values = c(1, 2),
  false_values = c(3, 4, 5, 6, 7, 8, 9),
  na_values = 97
) %>%
  pull(Dem_BChair_Age)
```

---

## 10. `label_likert()` --- Label Likert scales with descriptive strings

**The problem:** After (or instead of) collapsing Likert values to integers, you
need descriptive string labels for reporting. The extraction script has 8+
`across()` blocks doing this for different column groups, each with slightly
different labels.

`label_likert()` uses the same `mapping` vector approach as `collapse_likert()`,
then looks up the corresponding label. Values in `na_values` get `na_label`
(default `"Unsure"`) instead of `NA`.

### Before

```{r}
before <- survey %>%
  mutate(
    FinanceChng_Benefits = case_when(
      FinanceChng_Benefits %in% c(1, 2) ~ "Decrease",
      FinanceChng_Benefits == 3 ~ "No change",
      FinanceChng_Benefits %in% c(4, 5) ~ "Increase",
      FinanceChng_Benefits == 97 ~ "Unsure",
      .default = NA_character_
    )
  )

before %>% select(FinanceChng_Benefits)
```

### After

```{r}
after <- survey %>%
  label_likert("FinanceChng_Benefits",
               labels = c("Decrease", "No change", "Increase"))

after %>% select(FinanceChng_Benefits)
```

### Applying different labels to different column groups

The extraction script uses `label_likert()` for several groups of columns, each
with different label text. You can chain multiple calls:

```{r}
after <- survey %>%
  label_likert("FinanceChng_Benefits",
               labels = c("Decrease", "No change", "Increase")) %>%
  label_likert("FinanceChng_Salaries",
               labels = c("Decrease in salaries", "No change",
                          "Increase in salaries")) %>%
  label_likert("FinanceChng_TotExp",
               labels = c("Decrease in expenses", "No change",
                          "Increase in expenses"))

after %>% select(starts_with("FinanceChng"))
```

### Option: custom `na_label`

The default `na_label` is `"Unsure"`, but you can change it:

```{r}
survey %>%
  label_likert("FinanceChng_Benefits",
               labels = c("Decrease", "No change", "Increase"),
               na_label = "Not applicable") %>%
  select(FinanceChng_Benefits)
```

### Option: custom `mapping`

The default `mapping = c(1, 1, 2, 3, 3)` collapses a 5-point scale into 3
categories. You can provide a different mapping if your scale is different:

```{r}
# 3-point scale: keep each value as its own category
three_point <- data.frame(satisfaction = c(1, 2, 3, 97))

label_likert(three_point, "satisfaction",
             labels = c("Dissatisfied", "Neutral", "Satisfied"),
             mapping = c(1L, 2L, 3L)) %>%
  pull(satisfaction)
```

---

## Putting it all together

In a real NTIS cleaning script, you chain multiple ntistools calls into a single
pipeline. Here's a condensed example showing how the functions work together:

```{r}
cleaned <- survey %>%
  # Step 1: Combine related binary indicators into summary columns
  combine_binary(GeoAreas_Locally,
                 GeoAreas_Local, GeoAreas_MultipleLocal,
                 GeoAreas_RegionalWithin) %>%
  combine_binary(GeoAreas_Multistate,
                 GeoAreas_MultipleState, GeoAreas_RegionalAcross,
                 strict_na = TRUE) %>%

  # Step 2: Create any/count/all summaries for disruption indicators
  count_binary("LLDisruption", LLLost, LLDelay, LLStop) %>%

  # Step 3: Recode multi-level demographics to simple 0/1
  recode_binary(c("ProgDem_Children", "ProgDem_Elders")) %>%

  # Step 4: Remove sentinel values
  recode_sentinel("DonImportance") %>%

  # Step 5: Fill in valid-skip NAs using flag columns
  impute_from_flag("PplSrv_NumWait") %>%
  impute_from_flag(
    vars = c("Staff_RegVlntr_2023", "Staff_RegVlntr_2024"),
    flag_map = c(Staff_RegVlntr_2023 = "Staff_RegVlntr_NA",
                 Staff_RegVlntr_2024 = "Staff_RegVlntr_NA")
  ) %>%

  # Step 6: Propagate parent answers to children
  propagate_parent(c("Regulations_Federal", "Regulations_State"),
                   parent_var = "Regulations") %>%

  # Step 7: Zero out variables for ineligible respondents
  apply_filter(c("StaffVacancies", "BenefitsImpact"),
               condition = HaveStaff == 1) %>%

  # Step 8: Label binary columns for reporting
  label_binary(
    labels = list(
      GeoAreas_Locally = c("Serving locally", "Not serving locally"),
      FinanceChng_Reserves = c("Drew on reserves", "Did not draw"),
      PrgSrvc_Suspend = c("Paused services", "Did not pause")
    ),
    na_values = 97
  ) %>%
  label_binary(
    labels = list(
      ProgDem_Children = c("Serving children", "Not serving children"),
      ProgDem_Elders = c("Serving seniors", "Not serving seniors")
    )
  ) %>%

  # Step 9: Label Likert scales for reporting
  label_likert(c("FinanceChng_Benefits", "FinanceChng_TotExp"),
               labels = c("Decrease", "No change", "Increase")) %>%
  label_likert("FinanceChng_Salaries",
               labels = c("Decrease in wages", "No change",
                          "Increase in wages"))

glimpse(cleaned)
```

What used to be hundreds of lines of repetitive `case_when()` calls becomes a
single, readable pipeline. Each step has a clear name that tells you what it
does, and the options are explicit rather than buried in copy-pasted code.

---

## Quick reference

| Function | Input | Output | Key options |
|---|---|---|---|
| `combine_binary()` | 0/1 columns | Single 0/1 column | `strict_na` |
| `count_binary()` | 0/1 columns | `_any`, `_count`, `_all` columns | --- |
| `recode_binary()` | Multi-level numeric | 0/1 integer | `ones`, `zeros`, `na_values` |
| `recode_sentinel()` | Numeric with sentinels | Numeric with NA | `values` |
| `impute_from_flag()` | Numeric + flag column | Numeric (NAs filled) | `flag_suffix`, `flag_map`, `impute_value` |
| `collapse_likert()` | 5-point Likert | 3-category integer | `mapping`, `na_values` |
| `propagate_parent()` | Parent + child columns | Children updated | `values` |
| `apply_filter()` | Any columns + condition | Columns with NAs added | `condition` |
| `label_binary()` | 0/1 columns | Character columns | `true_values`, `false_values`, `na_values` |
| `label_likert()` | Likert numeric | Character columns | `mapping`, `labels`, `na_values`, `na_label` |
