---
title: "Weighted Survey Statistics with ntistools"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Weighted Survey Statistics with ntistools}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
has_survey <- requireNamespace("survey", quietly = TRUE)
has_srvyr  <- requireNamespace("srvyr",  quietly = TRUE)
has_emmeans <- requireNamespace("emmeans", quietly = TRUE)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  eval     = has_survey && has_srvyr
)
```

## Overview

After cleaning NTIS data with the recode / label / combine functions, the next
step in most analyses is *weighted inference*: confidence intervals, chi-square
tests of association, t-tests across two groups, and ANOVA across many groups.
The `survey` and `srvyr` packages provide the underlying machinery, but the
calls are verbose and the outputs are not tidy — making it tedious to run the
same test over many variables and stack the results into a single CSV.

ntistools provides five functions that wrap these patterns:

| Function | Wraps | Use when |
|---|---|---|
| `survey_ci()` | `srvyr::survey_mean(vartype = "ci")` | You want weighted means + CIs for many variables, optionally by a group. |
| `survey_chisq()` | `survey::svychisq()` | You want pairwise tests of association across many variable pairs. |
| `survey_ttest()` | `survey::svyttest()` | You want a t-test of each of several outcomes against a binary group. |
| `survey_anova()` | `survey::svyglm()` + `survey::regTermTest()`, optionally `emmeans` | You want a global F-test (and optional pairwise contrasts) for each outcome against a multi-level group. |
| `run_survey_stats()` | All of the above | You want to drive all four from spec data frames and write CSVs. |

All five return tidy data frames with one row per result, so they stack
cleanly with `dplyr::bind_rows()` or write straight to CSV.

> **Optional dependencies.** `survey`, `srvyr`, and `emmeans` are listed in
> `Suggests`. The functions load them lazily and emit a clear install hint if
> a package is missing — you only need to install what you use.

## Setup

```{r setup, message = FALSE}
library(ntistools)
library(dplyr)

set.seed(1)
n <- 200
svy <- data.frame(
  SizeStrata        = sample(c("Small", "Medium", "Large"), n, replace = TRUE),
  census_urban_area = sample(c("Urban", "Rural"), n, replace = TRUE),
  LLAnyDisruption   = rbinom(n, 1, 0.45),
  LLLost            = rbinom(n, 1, 0.30),
  LLLostAmt         = round(rlnorm(n, log(20000), 0.6)),
  weight_year6plus  = runif(n, 0.5, 3)
)
head(svy)
```

## `survey_ci()` --- weighted means with confidence intervals

Pass a character vector of variables and a weight column. Optionally pass a
`group_var` to break the estimates out by a categorical column.

```{r ci-national}
survey_ci(
  svy,
  vars   = c("LLAnyDisruption", "LLLost"),
  wt_var = "weight_year6plus"
)
```

Group by `SizeStrata` to get one row per (variable, group level):

```{r ci-grouped}
survey_ci(
  svy,
  vars      = c("LLAnyDisruption", "LLLost"),
  wt_var    = "weight_year6plus",
  group_var = "SizeStrata"
)
```

Use `level` to change the CI width (default `0.95`).

## `survey_chisq()` --- chi-square over variable pairs

Pass a two-column data frame describing the pairs to test. One row of output
per pair.

```{r chisq}
pairs <- data.frame(
  var1 = c("SizeStrata",        "census_urban_area"),
  var2 = c("LLAnyDisruption",   "LLAnyDisruption")
)
survey_chisq(svy, pairs, wt_var = "weight_year6plus")
```

The output columns are `var1`, `var2`, `statistic`, `df`, `p_value` — no
extra unnesting required.

## `survey_ttest()` --- t-tests across many outcomes

Pass a character vector of outcomes and a single binary grouping variable.
One row of output per outcome.

```{r ttest}
survey_ttest(
  svy,
  outcomes  = c("LLLostAmt", "LLAnyDisruption"),
  group_var = "census_urban_area",
  wt_var    = "weight_year6plus"
)
```

## `survey_anova()` --- ANOVA across many outcomes, optional pairwise

For categorical groups with more than two levels. The grouping variable is
coerced to a factor automatically.

```{r anova-overall}
survey_anova(
  svy,
  outcomes  = "LLLostAmt",
  group_var = "SizeStrata",
  wt_var    = "weight_year6plus"
)
```

The return value is a list. With `pairwise = TRUE`, it gains a `$pairwise`
element with `emmeans`-style contrasts (requires the `emmeans` package):

```{r anova-pairwise, eval = has_survey && has_srvyr && has_emmeans}
res <- survey_anova(
  svy,
  outcomes  = "LLLostAmt",
  group_var = "SizeStrata",
  wt_var    = "weight_year6plus",
  pairwise  = TRUE,
  adjust    = "bonferroni"
)
res$pairwise
```

## `run_survey_stats()` --- bulk-run everything and write CSVs

When you want to run dozens of these tests across one or more datasets,
describe the jobs as spec data frames and let `run_survey_stats()` drive them.
Each spec references a dataset by name from the `datasets` list.

```{r run-stats}
out <- run_survey_stats(
  datasets   = list(y6 = svy),

  ci_spec    = data.frame(
    dataset   = "y6",
    variable  = c("LLAnyDisruption", "LLLost"),
    wt_var    = "weight_year6plus",
    group_var = "SizeStrata"
  ),

  chisq_spec = data.frame(
    dataset = "y6",
    var1    = c("SizeStrata", "census_urban_area"),
    var2    = c("LLAnyDisruption", "LLLost"),
    wt_var  = "weight_year6plus"
  ),

  ttest_spec = data.frame(
    dataset   = "y6",
    outcome   = c("LLLostAmt"),
    group_var = "census_urban_area",
    wt_var    = "weight_year6plus"
  ),

  anova_spec = data.frame(
    dataset   = "y6",
    outcome   = "LLLostAmt",
    group_var = "SizeStrata",
    wt_var    = "weight_year6plus"
  )
)

names(out)
out$ci
```

Pass `output_dir = "stats_out"` to write one CSV per analysis type
(`ci.csv`, `chisq.csv`, `ttest.csv`, `anova_overall.csv`, and
`anova_pairwise.csv` when applicable). Pass `combined = TRUE` to write a
single `survey_stats.csv` with an `analysis` column tagging each row.

### Running across multiple datasets

The `datasets` argument is a *named list*, and each spec row picks the
dataset by name via the `dataset` column. To mix year-6 and year-5 jobs,
hand both data frames to `run_survey_stats()` and reference them in the
spec:

```r
run_survey_stats(
  datasets = list(
    y6 = nptrends_y6_clean,
    y5 = nptrends_y5_clean
  ),
  ci_spec = data.frame(
    dataset   = c("y6", "y6", "y5"),
    variable  = c("LLAnyDisruption", "LLLost", "LLAnyDisruption"),
    wt_var    = c("weight_year6plus", "weight_year6plus", "year5wt"),
    group_var = c("SizeStrata", "SizeStrata", NA)
  ),
  output_dir = "stats_out",
  combined   = TRUE
)
```

`group_var` is optional in `ci_spec`; leave it as `NA` for a national
(ungrouped) estimate.

## When to use which

- **Descriptive summaries** (proportions, means, medians per group, no CIs):
  use `calc_summarize()` / `summarize_by_groups()`. See
  `vignette("before-and-after")`.
- **Descriptive summaries with confidence intervals**: use `survey_ci()`.
- **Test of association between two categorical variables**: `survey_chisq()`.
- **Compare a continuous (or 0/1) outcome across two groups**: `survey_ttest()`.
- **Compare a continuous outcome across 3+ groups**: `survey_anova()`,
  with `pairwise = TRUE` if you want post-hoc contrasts.
- **A whole battery of the above at once**: `run_survey_stats()`.
